WHOOMP! (There It Is) \\ Rapid Bayesian position reconstruction for gravitational\nobreakdashes-wave transients

### 2014-09-17 ms.tex
marginpar to footnote
Rename section
Remove marginpar

\section{Theory of \ac{GW} detection and parameter estimation}

@@

The prevailing technique for detection of \acp{GW} from \acp{CBC} is to realize a \ac{MLE} from the likelihood in Equation~(\ref{eq:gaussian-likelihood}) and the signal model in Equation~(\ref{eq:signal-model}). Concretely, this results in a bank of matched filters, or the cross-correlation between a template waveform and the incoming data stream,
%
\begin{equation}
z_i(\tau_i;\bm\theta_\mathrm{in}) = \frac{1}{\sigma_i (\bm\theta_\mathrm{in})} \int_0^\infty \frac{h^*(\omega; \bm\theta_\mathrm{in}) y_i(\omega) e^{i \omega \tau_i}}{S_i(\omega)} \,d\omega.
\end{equation}
%
The \ac{ML} point estimates of the signal parameters, $\mathrm{MLE}(\mathbf{y}) = \{\{ \hat{\bm\theta}_i \}_i, \hat{\bm\theta}_\mathrm{in}\} = \{\left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i, \hat{\bm\theta}_\mathrm{in}\}$, are given by
%
\begin{eqnarray}
    \hat{\bm\theta}_\mathrm{in}, \{\hat\tau_i\}_i
        &=& \argmax_{\bm\theta_\mathrm{in}, \{\hat\tau_i\}_i}
        \sum_i \left| z_i\left(\tau_i;
        \bm\theta_\mathrm{in}\right) \right|^2, \\
    \hat\rho_i &=& \left| z_i\left(\hat\tau_i;
        \hat{\bm\theta}_\mathrm{in}\right) \right|, \\
    \hat\gamma_i &=& \arg z_i\left(\hat\tau_i;
        \hat{\bm\theta}_\mathrm{in}\right).
\end{eqnarray}
%
A detection candidate consists of $\{\left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i, \hat{\bm\theta}_\mathrm{in}\}$. There are various ways to characterize the significance of a detection candidate, but in Gaussian noise the combined SNR $\hat\rho_\mathrm{c}$ suffices,
%
\marginpar{What is the least that I can get away with saying about false alarm rate estimation and likelihood ratios?}%
%
\begin{equation}
    \hat\rho_\mathrm{c} = \sqrt{\sum_i {\hat\rho_i}^2}.
\end{equation}

@@

We can now write the Fisher matrix for the autocorrelation likelihood in a way that makes a comparison to the full signal model explicit. Define:
%
\begin{eqnarray*}
    \mathcal{I}_{\gamma\gamma} &=& \rho^2 \cdot \textproto{D}_{\gamma\gamma}(\rho) \\
    \mathcal{I}_{\tau\tau} &=& \rho^2 \overline{\omega^2} \cdot \textproto{D}_{\tau\tau}(\rho) \\
    \mathcal{I}_{\gamma\tau} &=& -\rho^2 \overline{\omega} \cdot \textproto{D}_{\gamma\tau}(\rho)
\end{eqnarray*}
%
Now, the $\textproto{D}_{ij}$ %
%
\marginpar{The Fish(er) factor. Go home, Leo.}%
%
contain the integrals from Equations~(\ref{eq:fisher-autocor-gamma-gamma}, \ref{eq:fisher-autocor-tau-tau}, \ref{eq:fisher-autocor-gamma-tau}) and encode the departure of the autocorrelation likelihood from the likelihood of the full data at low SNR. All of the $\textproto{D}_{ij}(\rho)$ are sigmoid-type functions that asymptotically approach 1 as $\rho \rightarrow \infty$. The transition SNR $\rho_\mathrm{crit}$ is largely the same for all three nontrivial matrix elements, and is determined by the time coincidence window $T$ and the signal bandwidth $\omega_\mathrm{rms}$.
%
\marginpar{The transition SNR is probably proportional to some power of $(\omega_\mathrm{rms} T)$.}

@@

\begin{acknowledgements}
\marginpar{This URL is not specific enough.}Source code for \ac{BAYESTAR} is available on the LIGO \acl{DASWG} web site at \url{http://www.lsc-group.phys.uwm.edu/daswg/projects/lalsuite.html}.

>>>>

\section{\ac{GW} detection and parameter estimation}

@@

The prevailing technique for detection of \acp{GW} from \acp{CBC} is to realize a \ac{MLE} from the likelihood in Equation~(\ref{eq:gaussian-likelihood}) and the signal model in Equation~(\ref{eq:signal-model}). Concretely, this results in a bank of matched filters, or the cross-correlation between a template waveform and the incoming data stream,
%
\begin{equation}
z_i(\tau_i;\bm\theta_\mathrm{in}) = \frac{1}{\sigma_i (\bm\theta_\mathrm{in})} \int_0^\infty \frac{h^*(\omega; \bm\theta_\mathrm{in}) y_i(\omega) e^{i \omega \tau_i}}{S_i(\omega)} \,d\omega.
\end{equation}
%
The \ac{ML} point estimates of the signal parameters, $\mathrm{MLE}(\mathbf{y}) = \{\{ \hat{\bm\theta}_i \}_i, \hat{\bm\theta}_\mathrm{in}\} = \{\left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i, \hat{\bm\theta}_\mathrm{in}\}$, are given by
%
\begin{eqnarray}
    \hat{\bm\theta}_\mathrm{in}, \{\hat\tau_i\}_i
        &=& \argmax_{\bm\theta_\mathrm{in}, \{\hat\tau_i\}_i}
        \sum_i \left| z_i\left(\tau_i;
        \bm\theta_\mathrm{in}\right) \right|^2, \\
    \hat\rho_i &=& \left| z_i\left(\hat\tau_i;
        \hat{\bm\theta}_\mathrm{in}\right) \right|, \\
    \hat\gamma_i &=& \arg z_i\left(\hat\tau_i;
        \hat{\bm\theta}_\mathrm{in}\right).
\end{eqnarray}
%
A detection candidate consists of $\{\left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i, \hat{\bm\theta}_\mathrm{in}\}$. There are various ways to characterize the significance of a detection candidate, but in Gaussian noise the combined SNR $\hat\rho_\mathrm{c}$ suffices,
%
\begin{equation}
    \hat\rho_\mathrm{c} = \sqrt{\sum_i {\hat\rho_i}^2}.
\end{equation}

@@

We can now write the Fisher matrix for the autocorrelation likelihood in a way that makes a comparison to the full signal model explicit. Define:
%
\begin{eqnarray*}
    \mathcal{I}_{\gamma\gamma} &=& \rho^2 \cdot \textproto{D}_{\gamma\gamma}(\rho) \\
    \mathcal{I}_{\tau\tau} &=& \rho^2 \overline{\omega^2} \cdot \textproto{D}_{\tau\tau}(\rho) \\
    \mathcal{I}_{\gamma\tau} &=& -\rho^2 \overline{\omega} \cdot \textproto{D}_{\gamma\tau}(\rho)
\end{eqnarray*}
%
Now, the $\textproto{D}_{ij}$ %
%
\footnote{The Fish(er) factor.}%
%
contain the integrals from Equations~(\ref{eq:fisher-autocor-gamma-gamma}, \ref{eq:fisher-autocor-tau-tau}, \ref{eq:fisher-autocor-gamma-tau}) and encode the departure of the autocorrelation likelihood from the likelihood of the full data at low SNR. All of the $\textproto{D}_{ij}(\rho)$ are sigmoid-type functions that asymptotically approach 1 as $\rho \rightarrow \infty$. The transition SNR $\rho_\mathrm{crit}$ is largely the same for all three nontrivial matrix elements, and is determined by the time coincidence window $T$ and the signal bandwidth $\omega_\mathrm{rms}$.
%
\marginpar{The transition SNR is probably proportional to some power of $(\omega_\mathrm{rms} T)$.}

@@

\begin{acknowledgements}
Source code for \ac{BAYESTAR} is available as part of LALInference, the open source LIGO/Virgo parameter estimation toolchain. See \url{http://www.lsc-group.phys.uwm.edu/daswg/projects/lalsuite.html} for further details.
